---
title: "beer"
author: "Gavin Fay"
date: "3/31/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(skimr)
library(broom)
```


## #TidyTuesday 2020

`TidyTuesday` is a weekly social data project in R.  
<https://github.com/rfordatascience/tidytuesday/blob/master/README.md>

This repository contains my work.  

### week 14, 2020-03-31, beer production

<https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-31/readme.md>

```{r getdata}

# Get the Data

brewing_materials <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv')
beer_taxed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_taxed.csv')
brewer_size <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewer_size.csv')
beer_states <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_states.csv')

# Or read in with tidytuesdayR package (https://github.com/thebioengineer/tidytuesdayR)
# PLEASE NOTE TO USE 2020 DATA YOU NEED TO USE tidytuesdayR version ? from GitHub

# Either ISO-8601 date or year/week works!

# Install via devtools::install_github("thebioengineer/tidytuesdayR")

#tuesdata <- tidytuesdayR::tt_load('2020-03-31')
#tuesdata <- tidytuesdayR::tt_load(2020, week = 14)

#brewing_materials <- tuesdata$brewing_materials
```

## Look at rate of increase in production by state
```{r bystate}
#skim(beer_states)
#unique(beer_states$type)
bs_use <- beer_states %>% 
  filter(type == "Kegs and Barrels") %>% 
  group_by(state) %>% 
  nest() %>% 
  #split(.$state) %>% 
  mutate(model = map(data, ~lm(log(barrels)~year, data = .x))) %>% 
         #summary = map(model, ~summary(,)),
  mutate(coefs = map(model, tidy, conf.int = TRUE)) %>% 
  unnest(coefs) %>% 
  filter(term == "year") %>% 
  I()
p1 <- bs_use %>% 
  ungroup() %>%
  #arrange(desc(estimate)) %>% 
  top_n(10, estimate) %>%
  mutate(state = fct_reorder(state, estimate)) %>% 
  ggplot() +
  aes(x = state, y = 100*estimate) +
  geom_point(col="#DF8D03", size =3) +
  geom_errorbar(aes(ymin=100*conf.low, ymax=100*conf.high), width=0,
                size = 1.2,
                col = "#DF8D03") +
  coord_flip() +
  labs(y = "annual % increase",
       x = "",
       title = "Which states are increasing beer production the fastest?",
       subtitle = "estimated annual rate of increase in kegs & barrels since 2008",
       caption = "\n@gavin_fay\n#TidyTuesday 2020 wk 14, data from https://www.ttb.gov/beer/statistics") +
  theme_minimal() +
  theme(
      plot.margin = margin(.75, .5, .75, .5, "cm"),
    plot.background = element_rect(fill = "grey15"),
    axis.text = element_text(color = "white", size = 12),
    axis.title.y = element_blank(), #text(size = 10),
    panel.grid = element_blank(),
    text = element_text(family = "Rockwell", color = "white"),
    plot.title.position = "plot") +
  NULL
p1
ggsave(filename = "beer_states.png", p1, height = 4.5, width = 8, dpi=300)


# p2 <- bs_use %>% 
#   ggplot() +
#   aes(x = estimate) +
#   geom_histogram(col="white") +
#   labs(y = "number of states", 
#        x = "% increase per yr") +
#   theme_minimal()
# 
# ggsave(filename = "beer_states2.png", p2, height = 6, dpi = 300)

```
      

## Look at cumulative production over time in New England
```{r}
# beer color panel
colpanel <- c("#FFF897","#FAE96F","#F6C101","#EC9D00","#DF8D03","#C96E12")

# cumulative production using purrr
cum_bs <- beer_states %>% 
  filter(type == "Kegs and Barrels") %>% 
  group_by(state) %>% 
  mutate(cumsum = accumulate(barrels, `+`)) %>% 
  I()
```

```{r}
p1 <- cum_bs %>% 
  filter(state %in% c("MA","NH","ME","VT","CT","RI")) %>% 
  ggplot() +
  aes(x = factor(year), y = barrels/1000000, group = state, fill = state) +
  #aes(x = factor(year), y = cumsum/1000000, group = state, fill = state) +
  geom_col(col = "white", position = "stack") +
  scale_fill_manual(values = colpanel) +
  theme_minimal() +
  theme(
      plot.margin = margin(.75, .5, .75, .5, "cm"),
    plot.background = element_rect(fill = "grey15"),
    axis.text = element_text(color = "white", size = 12),
    axis.title.x = element_text(size = 10),
    panel.grid = element_blank(),
    text = element_text(family = "Rockwell", color = "white"),
    plot.title.position = "plot",
    legend.position = c(0, 0.95), #"top",
    legend.justification='left') +
    guides(fill = guide_legend(nrow = 1)) +
  labs(x = "year",
       y = "barrels (millions)",
       fill = NULL,
       title = "New England beer production",
       subtitle = "kegs & barrels by state") +
  NULL


p2 <- cum_bs %>% 
  filter(state %in% c("MA","NH","ME","VT","CT","RI")) %>% 
  ggplot() +
  aes(x = factor(year), y = cumsum/1000000, group = state, fill = state) +
  geom_col(col = "white", position = "stack") +
  scale_fill_manual(values = colpanel) +
  theme_minimal() +
  theme(
      plot.margin = margin(.75, .5, .75, .5, "cm"),
    plot.background = element_rect(fill = "grey15"),
    axis.text = element_text(color = "white", size = 12),
    axis.title.x = element_text(size = 10),
    panel.grid = element_blank(),
    text = element_text(family = "Rockwell", color = "white"),
    plot.title.position = "plot",
    legend.position = c(0, 0.95), #"top",
    legend.justification='left') +
    guides(fill = guide_legend(nrow = 1)) +
  labs(x = "year",
       y = "barrels (millions)",
       fill = NULL,
       title = "New England beer production",
       subtitle = "cumulative kegs & barrels since 2008",
       caption = "\n@gavin_fay\n#TidyTuesday 2020 wk 14, data from https://www.ttb.gov/beer/statistics") +
  scale_y_continuous(labels=c(0, 2, 4, 6),breaks=c(0,2,4,6)) +
  NULL
p2
ggsave("cumuplot.png",p2,height = 4.5, width = 8, dpi=300)
ggsave("bothplots.png",p1/p2,height = 9, width = 8, dpi=300)

```



### things I learned and got stuck on...

* more practice with dplyr & `purrr`.  
* `accumulate` function from `purrr`.   
* using a custom color palette with `scale_discrete_manual`.  
* more practice with theme adjustments.  
* manual axis labels (to remove the pesky '8')  
* thanks to `@ekholm_e` for theme adjustment inspiration.  
* thanks to last week's meetup, I am LOVING ending pipes with `I()` & plots with `NULL`. Game-changer.    
