---
title: "broadway"
author: "Gavin Fay"
date: "4/28/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(skimr)
library(broom)
library(lubridate)
#install.packages('fishualize')
library(fishualize)
library(showtext)
library(sysfonts)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")

```


## #TidyTuesday 2020

`TidyTuesday` is a weekly social data project in R.  
<https://github.com/rfordatascience/tidytuesday/blob/master/README.md>

This repository contains my work.  

### week 18, 2020-04-28, Broadway shows

<https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-04-28/readme.md>

```{r getdata}
# Get the Data

tuesdata <- tidytuesdayR::tt_load(2020, week = 18)

grosses_raw <- tuesdata$grosses
synopses <- tuesdata$synopses
```

## look at multiple runs
### clean data
```{r}
grosses_raw <- tuesdata$grosses
grosses_fixed_missing <- grosses_raw %>%
  # Turn metrics into missing values if there were no shows
  # OR if metrics have a value of zero
  mutate_at(vars(weekly_gross:pct_capacity),
            ~ ifelse(performances + previews == 0 | . == 0, NA, .))

grosses_clean_temp <- grosses_fixed_missing %>%
  group_by(show) %>%
  arrange(week_ending) %>%
  mutate(run_number = cumsum(row_number() == 1 |
                               week_ending - lag(week_ending) > 90)) %>%
  group_by(show, run_number) %>%
  mutate(week_of_run = row_number()) %>%
  ungroup()

pre_1985_starts <- readr::read_csv("https://raw.githubusercontent.com/tacookson/data/master/broadway-grosses/pre-1985-starts.csv")

calculate_weeks_since_start <- function(x) {
  as.integer(pmax(1, difftime("1985-06-09", x, units = "weeks")))
}

pre_1985_starts_calculated <- grosses_clean_temp %>%
  group_by(show, run_number) %>%
  filter(min(week_ending) == "1985-06-09") %>%
  ungroup() %>%
  select(week_ending, show) %>%
  left_join(pre_1985_starts, by = "show") %>%
  group_by(show) %>%
  mutate(week_of_run_originals = calculate_weeks_since_start(start_date) + row_number()) %>%
  ungroup() %>%
  select(week_ending, show, week_of_run_originals)

grosses_clean <- grosses_clean_temp %>%
  left_join(pre_1985_starts_calculated, by = c("show", "week_ending")) %>%
  mutate(week_of_run = coalesce(week_of_run_originals, week_of_run)) %>%
  select(-week_of_run_originals)

cpi <- tuesdata$cpi %>%
  mutate(jan_2020_dollars = cpi[year_month == "2020-01-01"] / cpi)

real_grosses <- grosses_clean %>%
  mutate(year_month = floor_date(week_ending, unit = "month")) %>%
  left_join(cpi, by = "year_month") %>%
  mutate_at(
    vars(
      weekly_gross_overall,
      weekly_gross,
      potential_gross,
      avg_ticket_price,
      top_ticket_price
    ),
    ~ . * jan_2020_dollars
  ) %>%
  select(-year_month:-jan_2020_dollars)

```


<!-- ```{r} -->
<!-- grosses_fixed_missing %>% -->
<!--   filter(show == "Les Miserables") %>% -->
<!--   ggplot(aes(week_ending, seats_sold)) + -->
<!--   geom_col() + -->
<!--   scale_x_date(date_breaks = "5 years", date_labels = "%Y") + -->
<!--   labs(title = "Two-four-six-oh-what??", -->
<!--        subtitle = "Seats sold by week for Les Misérables") -->
<!-- ``` -->


## look at multuple runs
```{r}
numruns <- real_grosses %>% 
  group_by(show) %>% 
  summarize(numruns = n_distinct(run_number))
numruns
```

```{r}
#real_grosses
gross_sum <- real_grosses %>% 
  left_join(numruns) %>% 
  filter(numruns > 1) %>% 
  group_by(show, run_number) %>% 
  summarize(sum_gross = sum(weekly_gross),
            nweeks = length(weekly_gross),
            avg_gross = sum_gross/nweeks)
#gross_sum
```

```{r}
gross_sum %>%
  ggplot() +
  aes(x = factor(run_number), y = avg_gross) +
  geom_boxplot()
```

# set up a theme (much pain with the Broadway font)
add the broadway font
```{r}
font_add(family = "Broadway", "Broadway\ Regular/Broadway\ Regular.ttf")
showtext_auto()
font_add_google("Roboto", "Roboto")
```
create a raincloud theme
```{r}
library(extrafont)
raincloud_theme <- theme_minimal() +
  theme(
  plot.margin = margin(0.5, 0.2, 0.2, 0.2, "cm"),  
  axis.title.x = element_text(family = "Roboto", color = "white"),
  axis.title.y = element_text(family = "Roboto", color = "white"),
  panel.border = element_blank(),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_blank(),
  axis.line.x = element_line(colour = "grey15", size = 0.5, linetype = "solid"),
  axis.line.y = element_line(colour = "grey15", size = 0.5, linetype = "solid"),
  axis.text = element_text(color = "white", size = 12),
  plot.caption = element_text(family = "Roboto", color = "white"),
  plot.title = element_text(family = "Broadway", color = "white", size = 24), 
  plot.subtitle = element_text(family = "Roboto", color = "white"),
  plot.background = element_rect(fill = "grey15"))
```


raincloud plot of avg weekly gross by run number
```{r}
g <- 
  gross_sum %>% 
  ggplot() +
  aes(x = factor(run_number), y = avg_gross/1000000, fill = factor(run_number)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8) +
  geom_point(aes(y = avg_gross/1000000, color = factor(run_number)), 
             position = position_jitter(width = .15), size = .75, alpha = 0.6) +
  guides(fill = FALSE) +
  guides(color = FALSE) +
  coord_flip() + # flip or not?
  scale_color_fish(option = "Lepomis_megalotis", discrete = TRUE, end = 0.6) +
  scale_fill_fish(option = "Lepomis_megalotis", discrete = TRUE, end = 0.6) +
  labs(title = "The Show Must Go On",
       subtitle = "More runs, more money: revivals of Broadway shows make more money",
       caption = "\n@gavin_fay\n#TidyTuesday 2020 wk 18, data courtesy of @alexcookson",
       x = "Run Number",
       y = "Average Weekly Gross (millions)") +
  ggrepel::geom_label_repel(data = (arrange(gross_sum, desc(avg_gross)) %>%
                              ungroup() %>% slice(1)),
                            aes(x = factor(run_number),
                                y = avg_gross/1000000,
                                label=show),
                            nudge_x = 1,
                            col = "white") +
  raincloud_theme +
  NULL
g
```

## make an inset plot with  a histogram of the ratios of revival grosses 
```{r}
init <- gross_sum %>% 
  ungroup() %>%
  filter(run_number == 1) %>% 
  mutate(init = avg_gross) %>% 
  select(show, init)

bob <- gross_sum %>% 
  ungroup() %>% 
  filter(run_number != 1) %>% 
  left_join(init) %>% 
  mutate(ratio = avg_gross/init,
         over1 = ifelse(ratio>1,1,0))

col_use <- fish(n = 5, option = "Lepomis_megalotis", end = 1)[c(1,3)]
pp <- bob %>% 
  ggplot() +
  aes(x = ratio) +
  geom_histogram(col="grey15", fill = col_use[1]) +
  geom_vline(xintercept = 1, lty=2, size = 0.75, col = col_use[2]) +
  raincloud_theme +
  labs(x = "revival:initial",
       y = "") +
  xlim(0,5) +
  theme(axis.line.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 8)) +
  NULL
```

```{r}
g3 <- g + annotation_custom(ggplotGrob(pp), xmin = 3.5, xmax = 6.5, 
                       ymin = 1.4, ymax = 1.8)
ggsave("broadway.png",g3,height = 4.5, width = 8, dpi=300)
```
    

### things I learned and got stuck on...

* My first raincloud plot!  
* more practice with themes.  
* adding fonts from `extrafont`.     
* using color palette from `fishualize`.  
* practice with an inset plot.  
* ggrepel (thanks @mcsiple!)  
