---
title: "schrute"
author: "Gavin Fay"
date: "3/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("schrute")
library(tidyverse)
library(schrute)
```


## #TidyTuesday 2020

`TidyTuesday` is a weekly social data project in R.  
<https://github.com/rfordatascience/tidytuesday/blob/master/README.md>

This repository contains my work.  

### week 12, 2020-03-17, the Office

View data, & investigate which writers used the phrase "That's What She Said".  

```{r}
glimpse(theoffice)
# number of episodes for each writer
writers <- theoffice %>% 
  group_by(season, episode) %>% 
  slice(1) %>% 
  ungroup() %>% 
  group_by(writer) %>% 
  count() %>% 
  rename(n_episodes = n)
# number of uses of 'that's what she said'
twss_writers <- theoffice %>% 
  mutate(text = tolower(text)) %>% 
  slice(str_which(text, "what she sa")) %>% 
  group_by(writer) %>% 
  count() %>% 
  left_join(writers) %>% 
  mutate(freq = n/n_episodes)
twss_writers 
```

bar chart focusing on which writers did it the most.  

```{r}
my_subtitle <- expression(paste("which writers used `", bold("that's what she said"),"` the most?"))

twss_plot <- ggplot(twss_writers, 
                    aes(x=fct_reorder(writer,freq), y=freq)) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  theme(text=element_text(family="Courier New", size=14),
        plot.title = element_text(hjust = -1.95, size = 20,
                                  face = "bold"),
        plot.subtitle = element_text(hjust = 1.87),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  labs(title = "the office",
       y = "uses per episode",
       subtitle = my_subtitle,
              caption = "\n@gavin_fay\n#TidyTuesday 2020 wk 12, data from `schrute` pkg")


png('twss.png',width = 8,height=4.5,units = 'in',res=300)
twss_plot
dev.off()
```


### things I learned and got stuck on...

* first time using a little bit of stringr: `str_which()` instead of `grep()`.  
* additional fonts with `extrafont` pkg - ended up not using as Courier was a reasonable match for the Office font (American Typewriter medium).  
* practice with theme adjustments.  
* not happy with my solution for getting title / subtitle aligned with left of plot margin, `hjust = ` has different effects depending on plot size.  

